Order statistics and numerical optimization: the case of a single observation

### 2020-03-01 inst/cook.Rmd
fixes #39

```{r data}
H <- H +
    order_likelihood(c(5,2,3,4)) + # person 1 says 5 is the best, 4 the worst.
    order_likelihood(c(1,5,3,4)) + # person 2 says 1 is the best, 4 the worst.
    order_likelihood(c(1,4,2,5)) + # person 3 says 1 is the best, 5 the worst.
    order_likelihood(c(1,2,3,5)) + # person 4 says 1 is the best, 5 the worst.
    order_likelihood(c(2,1,3,4))   # person 5 says 2 is the best, 4 the worst.
H
```

@@

```{r}
library(partitions)
imH <- indep(maxp(H))
f <- function(o){loglik(imH,order_likelihood(o),log=FALSE)}
LL <- apply(perms(5),2,f)
length(LL)
sum(LL)
head(LL)
```

@@

```{r,beforetraining}
H <- hyper2(pnames=c(paste("p",1:5,sep=""),"t"))  # 't' the strength of training
# observations before training:
Hbefore <- H +
    order_likelihood(c(5,2,3,4,1)) + #  5 is the best, 4 the worst.
    order_likelihood(c(5,3,2,4,1)) + #  5 is the best, 1 the worst.
    order_likelihood(c(4,5,3,2,1)) + #  4 is the best, 1 the worst.
    order_likelihood(c(5,3,4,1,2)) + #  5 is the best, 2 the worst.
    order_likelihood(c(2,5,3,1,4))   #  2 is the best, 4 the worst.
Hbefore
```

>>>>

```{r data}
H <- H +
    rank_likelihood(c(5,2,3,4)) + # person 1 says 5 is the best, 4 the worst.
    rank_likelihood(c(1,5,3,4)) + # person 2 says 1 is the best, 4 the worst.
    rank_likelihood(c(1,4,2,5)) + # person 3 says 1 is the best, 5 the worst.
    rank_likelihood(c(1,2,3,5)) + # person 4 says 1 is the best, 5 the worst.
    rank_likelihood(c(2,1,3,4))   # person 5 says 2 is the best, 4 the worst.
H
```

@@

```{r}
library(partitions)
imH <- indep(maxp(H))
f <- function(o){loglik(imH,rank_likelihood(o),log=FALSE)}
LL <- apply(perms(5),2,f)
length(LL)
sum(LL)
head(LL)
```

@@

```{r,beforetraining}
H <- hyper2(pnames=c(paste("p",1:5,sep=""),"t"))  # 't' the strength of training
# observations before training:
Hbefore <- H +
    rank_likelihood(c(5,2,3,4,1)) + #  5 is the best, 4 the worst.
    rank_likelihood(c(5,3,2,4,1)) + #  5 is the best, 1 the worst.
    rank_likelihood(c(4,5,3,2,1)) + #  4 is the best, 1 the worst.
    rank_likelihood(c(5,3,4,1,2)) + #  5 is the best, 2 the worst.
    rank_likelihood(c(2,5,3,1,4))   #  2 is the best, 4 the worst.
Hbefore
```

### 2020-03-01 inst/formula1.Rmd
fixes #39

Each row is a driver and each column (after the first) a venue.  We
see that Hamilton, the first row, came second in Australia, first in
China, second in Bahrain, fourth in Russia, and so on.  In the first
column, we see the result from Australia (AUS) in which Hamilton came
second, Vettel first, Bottas third, and so on.  Notation used also
includes six different classes of no-score such as "Ret" for retired,
"WD" for withdrawn, and so on.  I now use bespoke R function
\code{ranktable2supp()}, that translates data of this type into a
log-likelihood.  The function is similar to, but slightly different
from, the analysis in `eurovision.R`.

```{r,applydata}
L2017 <- ranktable2supp(f2017)
head(L2017,3)
```

@@

```{r,fouryearsdata}
F1_2014 <- "formula1_2014.txt" %>% read.table(header=TRUE) %>% ranktable2supp
F1_2015 <- "formula1_2015.txt" %>% read.table(header=TRUE) %>% ranktable2supp
F1_2016 <- "formula1_2016.txt" %>% read.table(header=TRUE) %>% ranktable2supp
F1_2017 <- "formula1_2017.txt" %>% read.table(header=TRUE) %>% ranktable2supp
a <- list(F1_2014, F1_2015, F1_2016, F1_2017)
alldrivers <- all_pnames(a)

>>>>

Each row is a driver and each column (after the first) a venue.  We
see that Hamilton, the first row, came second in Australia, first in
China, second in Bahrain, fourth in Russia, and so on.  In the first
column, we see the result from Australia (AUS) in which Hamilton came
second, Vettel first, Bottas third, and so on.  Notation used also
includes six different classes of no-score such as "Ret" for retired,
"WD" for withdrawn, and so on.  I now use bespoke R function
\code{ordertable2supp()}, that translates data of this type into a
log-likelihood.  The function is similar to, but slightly different
from, the analysis in `eurovision.R`.

```{r,applydata}
L2017 <- ordertable2supp(f2017)
head(L2017,3)
```

@@

```{r,fouryearsdata}
F1_2014 <- "formula1_2014.txt" %>% read.table(header=TRUE) %>% ordertable2supp
F1_2015 <- "formula1_2015.txt" %>% read.table(header=TRUE) %>% ordertable2supp
F1_2016 <- "formula1_2016.txt" %>% read.table(header=TRUE) %>% ordertable2supp
F1_2017 <- "formula1_2017.txt" %>% read.table(header=TRUE) %>% ordertable2supp
a <- list(F1_2014, F1_2015, F1_2016, F1_2017)
alldrivers <- all_pnames(a)

### 2020-03-01 inst/loser.Rmd
more work
fixes #39

```{r}
n <- 6 # number of competitors
H <- hyper2(pnames=letters[1:6])
L <- ggol(H,letters[2:n],letters[1]) 
length(L)
L[[1]]
L[[2]]
```

@@

# Discussion

@@

## Repeated observations

@@

>>>>

```{r}
n <- 6 # number of competitors
H <- hyper2(pnames=letters[1:6])
L <- ggrl(H,letters[2:n],letters[1]) 
length(L)
L[[1]]
L[[2]]
```

@@

# Discussion: repeated observations

@@

@@

Note carefully that the observation considered here is "I came last".
This is different from the observation that "I did not win" which is
considered separately in `inst/notthewinner.Rmd`.

### 2020-03-01 inst/moto.Rmd
fixes #39

Each row is a driver (rider?) and each column (after the first) a
venue.  Takin the first row we see that Marquez came second in Qatar,
first in Argentina, retired in Austin, and so on.  In the first
column, we see the result from Qatar (QAT) in which Marquez came
second, Dovizioso first, Vinales third, and so on.  Notation used also
includes six different classes of no-score such as "Ret" for retired,
"WD" for withdrawn, and so on.  I now use bespoke R function
\code{ranktable2supp()}, that translates data of this type into a
log-likelihood.  The function is similar to, but slightly different
from, the analysis in `eurovision.R`.

@@

(object `motoGP_2019` is equivalent to
`ranktable2supp(motoGP_2019_table)`.  Above, we see a small part of
the Placket-Luce likelihood function for the 2019 results.  We can
find the maximum likelihood estimator by using the `maxp()` function:

>>>>

Each row is a driver (rider?) and each column (after the first) a
venue.  Takin the first row we see that Marquez came second in Qatar,
first in Argentina, retired in Austin, and so on.  In the first
column, we see the result from Qatar (QAT) in which Marquez came
second, Dovizioso first, Vinales third, and so on.  Notation used also
includes six different classes of no-score such as "Ret" for retired,
"WD" for withdrawn, and so on.  I now use bespoke R function
\code{ordertable2supp()}, that translates data of this type into a
log-likelihood.  The function is similar to, but slightly different
from, the analysis in `eurovision.R`.

@@

(object `motoGP_2019` is equivalent to
`ordertable2supp(motoGP_2019_table)`.  Above, we see a small part of
the Placket-Luce likelihood function for the 2019 results.  We can
find the maximum likelihood estimator by using the `maxp()` function:

### 2020-03-01 inst/notthewinner.Rmd
slight rewording
fixes #39

```{r}
(H <- hyper2(pnames=paste("p",1:5,sep="")) +
    order_likelihood(c(5,2,1,4)) + 
    order_likelihood(c(2,4,1)) + 
    order_likelihood(c(4,3,1)) + 
    order_likelihood(c(3,5)))
	```

@@

```{r,uninformativesupportfunction}
library("partitions")
H # as before, here for convenience
M <- perms(5)   # uninformative!
SL <- list()
for(i in seq_len(ncol(M))){
	SL[[i]] <- H + order_likelihood(M[,i,drop=TRUE])
}
SL <- as.suplist(SL)
```

@@

Now we make another observation ${\mathcal O}$ that is weakly
informative: person `1` did not win.  Package idiom is similar to the
perfectly uninformative observation considered above, except that we
remove every order that has person 1 winning:

```{r,slacalc}
M <- perms(5)
M <- M[, M[1,]!=1]  # remove any column with "1" in first place
SLa <- list()
for(i in seq_len(ncol(M))){
	SLa[[i]] <- H + order_likelihood(M[,i,drop=TRUE])
}
SLa <- as.suplist(SLa)
```

>>>>

```{r}
(H <- hyper2(pnames=paste("p",1:5,sep="")) +
    rank_likelihood(c(5,2,1,4)) + 
    rank_likelihood(c(2,4,1)) + 
    rank_likelihood(c(4,3,1)) + 
    rank_likelihood(c(3,5)))
	```

@@

```{r,uninformativesupportfunction}
library("partitions")
H # as before, here for convenience
M <- perms(5)   # uninformative!
SL <- list()
for(i in seq_len(ncol(M))){
	SL[[i]] <- H + rank_likelihood(M[,i,drop=TRUE])
}
SL <- as.suplist(SL)
```

@@

Now we make another observation ${\mathcal O}$ which is independent of
${\mathcal I}$: person `1` did not win.  It is clear that observation
${\mathcal O}$ is only very weakly informative.  Package idiom is
similar to the perfectly uninformative observation considered above,
except that we remove every order that has person 1 winning:

```{r,slacalc}
M <- perms(5)
M <- M[, M[1,]!=1]  # remove any column with "1" in first place
SLa <- list()
for(i in seq_len(ncol(M))){
	SLa[[i]] <- H + rank_likelihood(M[,i,drop=TRUE])
}
SLa <- as.suplist(SLa)
```

### 2020-03-01 inst/skating_analysis.Rmd
fixes #39

Object `skating_table` is structured so that each competitor is a
row, and each judge is a column.  Function `order_likelihood()`
considers each row to be a race [or a judge], so we need to take a
transpose.  The following R idiom is plausible but incorrect:

```{r}
skating_incorrect <- order_likelihood(t(skating_table))
```

@@

```{r}
skating <- order_likelihood(t(skating_rank))
pnames(skating) <- rownames(skating_table)
head(skating)
```

>>>>

Object `skating_table` is structured so that each competitor is a
row, and each judge is a column.  Function `rank_likelihood()`
considers each row to be a race [or a judge], so we need to take a
transpose.  The following R idiom is plausible but incorrect:

```{r}
skating_incorrect <- rank_likelihood(t(skating_table))
```

@@

```{r}
skating <- rank_likelihood(t(skating_rank))
pnames(skating) <- rownames(skating_table)
head(skating)
```
